% Chapter 2

\chapter{Materials and methods} % Main chapter title
\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter2}

% NEED MODIFICATION

\bfseries{Cell culture}\\
\normalfont HEK293 cells were cultured under standard cell culture conditions. In brief, cells were grown at \SI{37}{\celsius} and 5\% CO\textsubscript{2} in T-75 flasks (CELLSTAR) at ~ 10\% confluency in DMEM (life technologies) complemented with 10\% fetal calf serum (Pan-Biotech). Cells used for SILAC based experiments were cultured in SILAC DMEM (life technologies) containing glutamine (Glutamax, life technologies), pyruvate (life technologies), non-essential amino acids (life technologies) and 10\% dialyzed fetal calf serum (Pan-Biotech). The SILAC DMEM was supplemented with standard L-arginine (Arg0, Sigma-Aldrich) and L-lysine (Lys0, Sigma-Aldrich)("light") as in 5. Alternatively, Arg6 and Lys4 ("medium") or Arg10 and Lys8 ("heavy") were added in place of their light counterparts.\\
For the BioID experiments HEK cells were cultured in either light, medium or heavy SILAC medium until fully labeled. The cells were seeded in 15 cm plates with ~ 25\% density. Two plates were used for each condition. \\
For the endocytosis inhibition experiments HEK cells were cultured in a 12-well plate coated with poly-lysine. The cells were \\
For the degradation inhibition experiments HEK cells were cultured in a 6-well plate. The cells were
\\*
\\*
\bfseries{Western blot}\\
\normalfont The Pierce BCA Protein Assay Kit was used for the colorimetric detection and quantification of total protein obtained from extraction. Protein aliquots of 15 \si{\micro}g were heated with sample buffer at \SI{100}{\celsius} for 5 minutes, loaded and electrophoresed through pre-cast Mini-PROTEIN TGX gel and subsequently transferred to a PVDF membrane using semi-dry transfer system. After blocking in 5\% milk powder in TBST at room temperature for 1 h, the blot was cut into halves, each containing the same samples. The two blots were incubated overnight in 0.1\% Anti-MSI2 Antibody (ab50829) solution in TBST and 0.1\% Anti-actin antibody in TBST, respectively, at \SI{4}{\celsius}. After rinsing the blots 3 times for 8 minutes with TBST, the blots were incubated with 0.02\% Anti-mouse IgG VHH Single Domain Antibody (HRP) (ab191866) solution in TBST at room temperature for 1 h. The Msi2 and actin protein levels were determined by chemiluminescence.
\\*
\\*

\bfseries{BioID}\\
\normalfont b

\bfseries{Inhibitor treatment}\\
\\*
\bfseries{Fluorescence microscopy}\\
\normalfont a
\\*



%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
